---
title: "ANLY 512 - Problem Set 2"
subtitle: "Anscombe's quartet"
author: "Xiaowei Wan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Objectives

The objectives of this problem set is to orient you to a number of activities in `R`.  And to conduct a thoughtful exercise in appreciating the importance of data visualization.  For each question create a code chunk or text response that completes/answers the activity or question requested. Finally, upon completion name your final output `.html` file as: `YourName_ANLY512-Section-Year-Semester.html` and upload it to the "Problem Set 2" assignmenet on Moodle.



## Questions

1. Anscombes quartet is a set of 4 $x,y$ data sets that were published by *Francis Anscombe* in a 1973 paper [*Graphs in statistical analysis*](https://www.jstor.org/stable/2682899?seq=1#page_scan_tab_contents).  For this first question load the `anscombe` data that is part of the `library(datasets)` in `R`.  And assign that data to a new object called `data`.

```{r load dataset}
data <- anscombe
```

2. Summarise the data by calculating the mean, variance, for each column and the correlation between each pair (eg. x1 and y1, x2 and y2, etc) (Hint: use the `fBasics()` package!)

```{r summarise anscombe data, warning=FALSE, message=FALSE}
# assign variables x1, x2, x3, x4, y1, y2, y3, y4
x1 <- data$x1
x2 <- data$x2
x3 <- data$x3
x4 <- data$x4
y1 <- data$y1
y2 <- data$y2
y3 <- data$y3
y4 <- data$y4
# mean and variance of x1
mean(x1)
var(x1)
# mean and variance of x2
mean(x2)
var(x2)
# mean and variance of x3
mean(x3)
var(x3)
# mean and variance of x4
mean(x4)
var(x4)

# mean and variance of y1
mean(y1)
var(y1)
# mean and variance of y2
mean(y2)
var(y2)
# mean and variance of y3
mean(y3)
var(y3)
# mean and variance of y4
mean(y4)
var(y4)

library("fBasics")
correlationTest(x1, y1)
correlationTest(x2, y2)
correlationTest(x3, y3)
correlationTest(x4, y4)
```

3. Create scatter plots for each $x, y$ pair of data.
```{r scatter data plots}
plot(x1, y1, main="Scatter plots between x1, y1")
plot(x2, y2, main="Scatter plots between x2, y2")
plot(x3, y3, main="Scatter plots between x3, y3")
plot(x4, y4, main="Scatter plots between x4, y4")
```

4. Now change the symbols on the scatter plots to solid circles and plot them together as a 4 panel graphic

```{r scatter plots with solid circles}
par(mfrow=c(2,2))
plot(x1, y1, main="Scatter plots between x1, y1", pch=16, col="red")
plot(x2, y2, main="Scatter plots between x2, y2", pch=16, col="blue")
plot(x3, y3, main="Scatter plots between x3, y3", pch=16, col="green")
plot(x4, y4, main="Scatter plots between x4, y4", pch=16, col="black")
```

5. Now fit a linear model to each data set using the `lm()` function.
```{r, anscombe linear model}
# linear model between x1 ~ y1
lm_1 <- lm(y1 ~ x1)
summary(lm_1)
# linear model between x2 ~ y2
lm_2 <- lm(y2 ~ x2)
summary(lm_2)
# linear model between x3 ~ y3
lm_3 <- lm(y3 ~ x3)
summary(lm_3)
# linear model between x4 ~ y4
lm_4 <- lm(y4 ~ x4)
summary(lm_4)
```

6. Now combine the last two tasks.  Create a four panel scatter plot matrix that has both the data points and the regression lines. (hint: the model objects will carry over chunks!)
```{r, scatter plots and regression lines}
par(mfrow= c(2,2))

plot(x1, y1, main = "Scatter Plot between x1,y1", pch = 20, col="red")
abline(lm_1, col = "blue")

plot(x2, y2, main = "Scatter Plot between  x2,y2", pch = 20, col="red")
abline(lm_2, col = "blue")

plot(x3, y3, main = "Scatter Plot between  x3,y3", pch = 20, col="red")
abline(lm_3, col = "blue")

plot(x4, y4, main = "Scatter Plot between  x4,y4", pch = 20, col="red")
abline(lm_4, col = "blue")
```


7.  Now compare the model fits for each model object.
```{r, results="asis"}
anova(lm_1)
anova(lm_2)
anova(lm_3)
anova(lm_4)
```


8. In text, summarize the lesson of Anscombe's Quartet and what it says about the value of data visualization.

Summarize:

- Anscombe’s Quartet consists of four data set of similar kind of data. 
- Each data set consists of eleven pairs of values for x and y. 
- The summary statistics are almost similar for each column values and correlations.
- The four data set appear to be identical.
- Plots tells that each data has different story.

Concludes:

* Summary statistics may not tell us all stories about the data.
* Graphical representation or visualization of data can give us a more clear picture about data.
* Data visualization in form of graphical datasets is important.







